---
title: "Data Import"
author: "Maliha Safdar"
date: "2025-09-16"
output: git_document
---

```{r}
library(tidyverse)
library(readxl)
library(haven)
```

Let's import a dataset.

```{r}
litters_df = 
  read_csv("data_import_examples/FAS_litters.csv")
```

Look at your data!!

```{r}
names(litters_df)
```

Update the names in "litters_df"

```{r}
litters_df = janitor::clean_names(litters_df)
```

Look at data!!

```{r}
names(litters_df)
```
Sometimes skimming data is neat? It helps to neatly summarize data.

```{r}
skimr::skim(litters_df)
```

```{r}
litters_df = 
    read_csv("data_import_examples/FAS_litters.csv",
    skip = 10, col_names = FALSE)
```
```{r}
head(litters_df)
```
```{r}
view()
```



Let's import the second dataset FAS_pups.

```{r}
pups_df = 
  read_csv("data_import_examples/FAS_pups.csv")
```
```{r}
pups_df = 
  read_csv("data_import_examples/FAS_pups.csv"   
  skip = 3
  na = c("NA",".""")
  )

pups_df = janitor:: clean_names(pups_df)
  )
```

```{r}
litters_df
```



Okay what about excel

CSVs are really great but sometimes you get an excel file. 


```{r}
mlb_df = 
  read_excel("data_import_examples/mlb11.xlsx")
```

Import LotR Words
```{r}
fort_df = 
  read_excel("data_import_examples/LotRWords.xlsx" , range = "B3:D6"")
```


SAS????

Import the PULSE data

```{r}
pulse_df = 
  read_sas("data_import_examples/public_pulse_data.sas7bdat")

pulse_df = 
  janitor :: clean_names(pulse_df)
```

Why do I have read.csv so much??

(after eval = FALSE he pressed knit)
```{r, eval = FALSE}
litters_df_base = 
  read.csv("data_import_examples/FAS_litters.csv")
```



What about data exporting?
it saves the file in your data_import_examples folder when use the write_csv code
```{r}
write_csv(fortr_df, "data_import_examples/fotr_df.csv")
```

He knit after this then commit the three files



