---
title: "Data_wrangling_i"
date: "09.16.2025"
output: github_document
---

```{r setup} 
library(tidyverse)
```

## Read some data, but first import it, and make variable names all lower case.

```{r}
litters_df = read_csv(file = "./data_import_examples/FAS_litters.csv")
```

```{r clean up variable names after importing data}
names(litters_df)

## use the janitor :: clean_names() function to put the variable names into snake lower case

litters_df = janitor::clean_names(litters_df)
names(litters_df)
```

## Importing the pups_df, and making all variables lower case.

```{r}
pups_df = read_csv(file = "./data_import_examples/FAS_pups.csv")

names(pups_df)

## use the janitor :: clean_names() function to put the variable names into snake lower case

litters_df = janitor::clean_names(pups_df)
names(pups_df)
```


### Looking at the data, and the head and tail.
```{r}
litters_df

head(litters_df, 5)

tail(litters_df, 5)
```


### Using the skimr function to quickly look at data.

```{r}
skimr::skim(litters_df)
```


### Arguments to read

```{r}
litters_df = 
    read_csv(file = "./data_import_examples/FAS_litters.csv",
    skip = 10, col_names = FALSE)

head(litters_df)
```


### Excluding the missing vaulues

```{r}
litters_df = 
    read_csv(
        file = "./data_import_examples/FAS_litters.csv",
        na = c(".", "NA", ""))
head(litters_df)
```

### Parsing columns

`In some cases, you’ll want to give explicit column specifications. This is done using the cols function, where you can give each column a type.`

```{r}
litters_df = 
    read_csv(file = "./data_import_examples/FAS_litters.csv",
        na = c(".", "NA", ""),
    col_types = cols(
      Group = col_character(),
      `Litter Number` = col_character(),
      `GD0 weight` = col_double(),
      `GD18 weight` = col_double(),
      `GD of Birth` = col_integer(),
      `Pups born alive` = col_integer(),
      `Pups dead @ birth` = col_integer(),
      `Pups survive` = col_integer()
    )
  )

tail(litters_df)
```

`Note that you don’t have to specify the variable type for every column, and can only focus on ones that are difficult:`

```{r}
litters_df = 
    read_csv(file = "./data_import_examples/FAS_litters.csv",
        na = c(".", "NA", ""),
    col_types = cols(
      Group = col_factor()
    )
)

head(litters_df)
```


#### Learning Assessment 1 

```{r}
pups_df = 
    read_csv(file = "./data_import_examples/FAS_pups.csv",
    skip = 3, col_names = FALSE)

head(pups_df)
```
 
 `excluding missing values.`
 
```{r}
pups_df = 
    read_csv(
        file = "./data_import_examples/FAS_pups.csv",
        skip = 3,
        na = c(".", "NA", ""))
head(pups_df)
```
 
 `parsing columns`
 
```{r}
pups_df = 
    read_csv(file = "./data_import_examples/FAS_pups.csv",
        skip = 3,
        na = c(".", "NA", ""),
    col_types = cols(
      `...1` = col_factor(),
      `1 = male, 2 = female` = col_double(),
      `...3` = col_double(),
      `...4` = col_double(),
      `...5` = col_double(),
      `...6` = col_double(),
      
    )
  )

skimr::skim(pups_df)
```
 
 `This works as well`
 
```{r}
pups_df = 
    read_csv("./data_import_examples/FAS_pups.csv",
        na = c(".", "NA"), col_types = "fddddd")

skimr::skim(pups_df)
```
 
 

### Other file formats


 